---
title: "Lab 6"
author: "Aksel Can Sozudogru"
date: "2024-06-28"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Task 1

```{r}
getwd()
```

## Task 2

```{r}
sample = runif(10, 0, 5)

uniMean = (0+5)/2
uniMean

uniVar = ((5-0)^2)/12
uniVar

xbar = mean(sample)
xbar

s2 = var(sample)
s2
```

```{r}
myclt=function(n,iter){
  y=runif(n*iter,0,5) # A
  data=matrix(y,nr=n,nc=iter,byrow=TRUE) #B
  sm=apply(data,2,sum) #C
  hist(sm)
  sm
}
w=myclt(n=10,iter=10000) #D
```

#### A is for generating random with a sample size of n*iter, lower limit of 0 and an upper limit of 5, and storing it in y. B is for matrix creation, which is called data and filling it up with the values stored in y. C sums the columns on the data matrix and stores it in a vector called sum. D calls the function and storing it in w.


```{r}
mean(w)
var(w)
```


```{r}
myclt=function(n,iter){
  y=runif(n*iter,0,5) # A
  data=matrix(y,nr=n,nc=iter,byrow=TRUE) #B
  m=apply(data,2, mean) #C
  hist(m)
  m
}
w=myclt(n=10,iter=10000) #D
```


```{r}
mean(w)
var(w)
```

## Task 3



```{r}
mycltu=function(n,iter,a=0,b=10){
## r-random sample from the uniform
y=runif(n*iter,a,b)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax
## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), main=paste("Histogram of sample mean",
"\n", "sample size= ",n,sep=""),xlab="Sample mean")
## add a density curve made from the sample distribution
lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=(a+b)/2,sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
## Add the density from which the samples were taken
curve(dunif(x,a,b),add=TRUE,lwd=4)

}

for (n in c(1, 2, 3, 5, 10, 30)) {
  mycltu(n = n, iter = 10000)
}
```

#### I conclude that our data distribution normalizes when n = 3, and our mean stays the same at all given sample sizes, supporting CLT.



## Task 4

```{r}
mycltb=function(n, iter = 10000, p = 0.3){

## r-random sample from the Binomial
y=rbinom(n*iter,size=n,prob=p)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Now we can make the histogram
## freq=FALSE means take a density
hist(w,freq=FALSE,  ylim=c(0,ymax),
main=paste("Histogram of sample mean","\n", "sample size= ",n, ", p= ",p,sep=""), xlab="Sample mean")
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=n*p,sd=sqrt(p*(1-p))),add=TRUE,col="Red",lty=2,lwd=3) 

}

for (n in c(4,5,10,20)) {
  mycltb(n = n)
}
```

```{r}
for (n in c(4,5,10,20)) {
  mycltb(n = n, p = 0.7)
}
```

```{r}
for (n in c(4,5,10,20)) {
  mycltb(n = n, p = 0.5)
}
```

#### CLT holds true for our distribution. When size is 4 some anomalies occur, but the rest is normal distribution.







## Task 5

```{r}
mycltp=function(n,iter = 10000,lambda=4){

## r-random sample from the Poisson
y=rpois(n*iter,lambda=lambda)
## Place these numbers into a matrix
## The columns will correspond to the iteration and the rows will equal the sample size n
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
## apply the function mean to the columns (2) of the matrix
## these are placed in a vector w
w=apply(data,2,mean)
## We will make a histogram of the values in w
## How high should we make y axis?
## All the values used to make a histogram are placed in param (nothing is plotted yet)
param=hist(w,plot=FALSE)
## Since the histogram will be a density plot we will find the max density

ymax=max(param$density)
## To be on the safe side we will add 10% more to this
ymax=1.1*ymax

## Make a suitable layout for graphing
layout(matrix(c(1,1,2,3),nr=2,nc=2, byrow=TRUE))

## Now we can make the histogram
hist(w,freq=FALSE,  ylim=c(0,ymax), col=rainbow(max(w)),
main=paste("Histogram of sample mean","\n", "sample size= ",n," iter=",iter," lambda=",lambda,sep=""), xlab="Sample mean")
## add a density curve made from the sample distribution
#lines(density(w),col="Blue",lwd=3) # add a density plot
## Add a theoretical normal curve 
curve(dnorm(x,mean=lambda,sd=sqrt(lambda/n)),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve

# Now make a new plot
# Since y is discrete we should use a barplot
barplot(table(y)/(n*iter),col=rainbow(max(y)), main="Barplot of sampled y", ylab ="Rel. Freq",xlab="y" )
x=0:max(y)
plot(x,dpois(x,lambda=lambda),type="h",lwd=5,col=rainbow(max(y)),
main="Probability function for Poisson", ylab="Probability",xlab="y")
}

for (n in c(2, 3, 5, 10, 20)) {
  mycltp(n = n)
}
```




```{r}
for (n in c(2, 3, 5, 10, 20)) {
  mycltp(n = n, lambda = 10)
}
```

<video width="320" height="240" controls>
  <source src="C:\Users\aksel\Videos\Captures\RStudio 2024-06-27 23-27-49.mp4" type="video/mp4">
Your browser does not support the video tag.
</video>

